---
title: "p8105_hw2_nl2835"
author: "Nancy Le"
date: 09-28-2023
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)

```

### Problem 1: 

We import and clean the FiveThirtyEight `pols_month_df` data, which gives the number of national politicians who are democratic or republican. Some values of `prez_gop` are `2` - we code them as `gop` in the new `prez` variable. 

```{r clean_538_pols}

month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_month_df = 
  read_csv(file = "./fivethirtyeight_datasets/pols-month.csv") |> 
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |> 
  mutate(
    prez = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 

```



We import and clean the FiveThirtyEight `SNP_df` data, which gives the monthly closing values of the S&P 500 by year. 

```{r clean_528_snp}
SNP_df = 
  read_csv(file = "./fivethirtyeight_datasets/snp.csv") |> 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) |> 
  mutate(month = month.name[month]) |> 
  arrange(year, month) |> 
  select(year, month, close)
```


We clean the `unemployment_df` data to merge with `pols_month_df` and `SNP_df` datasets.


```{r clean_538_unemp}
unemployment_df = 
  read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") |> 
  rename(year = Year) |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _,  y = month_df) |> 
  select(year, month, unemployment)
```

Merging 3 datasets

```{r merge 538}
fivethirtyeight_data = 
  left_join(pols_month_df, SNP_df) |> 
  left_join(x = _, y = unemployment_df)


str(fivethirtyeight_data)
```

The `unemployment_df` dataset described unemployment by month and year, the `SNP_df` dataset described closing values of the S&P 500 by month and year, and the `pols_month_df` dataset described the count of elected officials by political party, by month and year. 

The `unemployment_df` dataset has `r nrow(unemployment_df)` observations and `r ncol(unemployment_df` variables. The years range from `r range(unemployment_df$year)[1]` to `r range(unemployment_df$year)[2]`. 

It tells us xx


The `SNP_df` dataset has `r nrow(SNP_df)` observations and `r ncol(SNP_df` variables. The years range from `r range(SNP_df$year)[1]` to `r range(SNP_df$year)[2]`. 

It tells us xx



The `pols_month_df` dataset has `r nrow(pols_month_df)` observations and `r ncol(pols_month_df` variables. The years range from `r range(pols_month_df$year)[1]` to `r range(pols_month_df$year)[2]`. 

It tells us xx


The full FiveThirtyEight dataset has `r nrow(fivethirtyeight_data)` observations and `r ncol(fivethiryteight_data` variables. The years range from `r range(fivethirtyeight_data$year)[1]` to `r range(fivethirtyeight_data$year)[2]`.




Problem 2:

```{r}
trashwheel_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 1,
        range = "A2:N549") |> 
  mutate(sheet = "mister trashwheel") |> 
  janitor::clean_names()
```


```{r}
trashwheel_df = trashwheel_df |> 
  mutate(homes_powered_v2 = (weight_tons * 500) / 30) |> 
  select(-homes_powered)
```


```{r}
trashwheel_df |> 
  select(dumpster, year, month, everything())
```

```{r}
trashwheel_df[["year"]] <-
  as.numeric(trashwheel_df[["year"]])

```


```{r}
proftrashwheel_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 2,
        range = "A2:M96") |> 
  mutate(sheet = "professor trashwheel") |> 
  janitor::clean_names()
```

```{r}
proftrashwheel_df = proftrashwheel_df |> 
  mutate(homes_powered_v2 = (weight_tons * 500) / 30) |> 
  select(-homes_powered)
```

Imported, cleaned, and organized Professor Trash Wheel data.

```{r}
gwynnda_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 4,
        range = "A2:K108") |> 
  mutate(sheet = "gwynnda") |> 
  janitor::clean_names()
```

```{r}
gwynnda_df = gwynnda_df |> 
  mutate(homes_powered_v2 = (weight_tons * 500) / 30) |> 
  select(-homes_powered)
```

Imported, cleaned, and organized Gwynnda data.



```{r}
trashwheel_collection_data =  
     full_join(gwynnda_df, proftrashwheel_df, by = c("dumpster",
              "month", "year", "date", "weight_tons", "volume_cubic_yards", 
              "plastic_bottles", "polystyrene", "cigarette_butts",
              "homes_powered_v2", "sheet"))

```



```{r}
trashwheel_collection_data_total = 
     full_join(trashwheel_collection_data, trashwheel_df, by = c("dumpster",
              "month", "year", "date", "weight_tons", "volume_cubic_yards", 
              "plastic_bottles", "polystyrene", "cigarette_butts", 
              "sheet", "glass_bottles", "grocery_bags", "homes_powered_v2",
              "chip_bags"))

```


This combined data set has ```r nrow(trashwheel_collection_data_total)``` and
```r ncol(trashwheel_collection_data_total)``` columns. The variables are: dumpster #, month, year, date, weight in tons, volume in cubic yards, plastic bottles, polystyrene, cigarette butts, plastic bags, which sheet the data is from, homes powered, glass bottles, grocery bags, chip bags, and sports balls. 

```{r}
ptwtons = trashwheel_collection_data_total |> 
  filter(sheet=="professor trashwheel") |> 
  pull(weight_tons) |> 
  sum() 

```


The total weight of trash collected by Professor Trash Wheel was ```r ptwtons``` tons.


```{r}
totalcigsgwyn = trashwheel_collection_data_total |> 
  filter(sheet=="gwynnda", year=="2021", month=="July") |> 
  pull(cigarette_butts) |> 
  sum() 
```


The total number of cigarette butts collected by Gwynnda in July of 2021 was ```r totalcigsgwyn```.
 



Problem #3






