---
title: "p8105_hw2_nl2835"
author: "Nancy Le"
date: 09-28-2023
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)

```

### Problem 1: 

We import and clean the FiveThirtyEight `pols_month_df` data, which gives the number of national politicians who are democratic or republican. Some values of `prez_gop` are `2` - we code them as `gop` in the new `prez` variable. 

```{r clean_538_pols}

month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_month_df = 
  read_csv(file = "./fivethirtyeight_datasets/pols-month.csv") |> 
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |> 
  mutate(
    prez = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 

```



We import and clean the FiveThirtyEight `SNP_df` data, which gives the monthly closing values of the S&P 500 by year. 

```{r clean_528_snp}
SNP_df = 
  read_csv(file = "./fivethirtyeight_datasets/snp.csv") |> 
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) |> 
  mutate(month = month.name[month]) |> 
  arrange(year, month) |> 
  select(year, month, close)
```


We clean the `unemployment_df` data to merge with `pols_month_df` and `SNP_df` datasets.


```{r clean_538_unemp}
unemployment_df = 
  read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") |> 
  rename(year = Year) |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _,  y = month_df) |> 
  select(year, month, unemployment)
```

Merging 3 datasets

```{r merge 538}
fivethirtyeight_data = 
  left_join(pols_month_df, SNP_df) |> 
  left_join(x = _, y = unemployment_df)


str(fivethirtyeight_data)
```



###Problem 2:

```{r}
trashwheel_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 1,
        range = "A2:N549") |> 
  janitor::clean_names() |> 
  mutate(
    sheet = "mister trashwheel", homes_powered_v2 = (weight_tons * 500) / 
      30) |> 
  select(dumpster, year, month, everything(), -homes_powered)
```



```{r clean prof trashwheel names}
proftrashwheel_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 2,
        range = "A2:M96") |> 
  janitor::clean_names()
  

proftrashwheel_df = proftrashwheel_df |> 
  mutate(
    homes_powered_v2 = (weight_tons * 500) / 30) |> 
  mutate(sheet = "professor trashwheel") |> 
  select(dumpster, year, month, everything(), -homes_powered)


```


Imported, cleaned, and organized Professor Trash Wheel data.

```{r}
gwynnda_df = 
  readxl::read_excel("./mr_trashwheel/trashwheel_collection.xlsx", 4,
        range = "A2:K108") |> 
  mutate(sheet = "gwynnda") |> 
  janitor::clean_names()

gwynnda_df = gwynnda_df |> 
  mutate(homes_powered_v2 = (weight_tons * 500) / 30) |> 
  select(dumpster, year, month, everything(), -homes_powered)
```

Imported, cleaned, and organized Gwynnda data.

```{r}
proftrashwheel_df <- proftrashwheel_df |>   
    mutate(year = as.numeric(year))


trashwheel_df <- trashwheel_df |>   
    mutate(year = as.numeric(year))

```


```{r merge_trash}
trashwheel_collection = 
  left_join(gwynnda_df, proftrashwheel_df) |>
  left_join(x = _, y = trashwheel_df)

str(trashwheel_collection)
```

```{r}
trashwheel_collection_data =  
     full_join(gwynnda_df, proftrashwheel_df, by = c("dumpster",
              "month", "year", "date", "weight_tons", "volume_cubic_yards", 
              "plastic_bottles", "polystyrene", "cigarette_butts",
              "homes_powered_v2", "sheet"))

```



```{r}
trashwheel_collection_data_total = 
     full_join(trashwheel_collection_data, trashwheel_df, by = c("dumpster",
              "month", "year", "date", "weight_tons", "volume_cubic_yards", 
              "plastic_bottles", "polystyrene", "cigarette_butts", 
              "sheet", "glass_bottles", "grocery_bags", "homes_powered_v2",
              "chip_bags"))

```


This combined data set has ```r nrow(trashwheel_collection_data_total)``` and
```r ncol(trashwheel_collection_data_total)``` columns. The variables are: dumpster #, month, year, date, weight in tons, volume in cubic yards, plastic bottles, polystyrene, cigarette butts, plastic bags, which sheet the data is from, homes powered, glass bottles, grocery bags, chip bags, and sports balls. 

```{r}
ptwtons = trashwheel_collection_data_total |> 
  filter(sheet=="professor trashwheel") |> 
  pull(weight_tons) |> 
  sum() 

```


The total weight of trash collected by Professor Trash Wheel was ```r ptwtons``` tons.


```{r}
totalcigsgwyn = trashwheel_collection_data_total |> 
  filter(sheet=="gwynnda", year=="2021", month=="July") |> 
  pull(cigarette_butts) |> 
  sum() 
```


The total number of cigarette butts collected by Gwynnda in July of 2021 was ```r totalcigsgwyn```.
 



Problem #3

```{r import MCI baseline data}
baseline_df = 
  read_csv("./data_mci/mci_baseline.csv") |> 
  janitor::clean_names() 
```

```{r rename variables}
names(baseline_df)

baseline_df <- as_tibble(baseline_df) |>  
  rename(ID = x1,
        age_baseline = age_at_the_study_baseline,
         sex = x1_male_0_female, 
         education_years = years_of_education,
         carrier_status = x1_apoe4_carrier_0_apoe4_non_carrier,
         onset_age = age_at_the_onset_of_mci_missing_if_a_subject_remains_mci_free_during_the_follow_up_period) |> 
  mutate(
    as.character(x1_male_0_female, carrier_status)
  )

  
```


```{r import MCI amyloid data}
amyloid_df = 
  read_csv("./data_mci/mci_amyloid.csv") |> 
  janitor::clean_names() 
```



